---
title: "Assignment 10 - Relationship Knowledge CFA"
output: html_notebook
---


```{r}
library(haven)
library(psych)
library(tidyverse)
library(lavaan)
#library(semPlot)
library(rowr)
```


### Read in data
```{r}
dim_rel_scaled = read.csv('dim_rel_scaled.csv', row.names=1)

names(dim_rel_scaled)
```


## Return the covariance matrix for the lower triangle
```{r}
m1_cov_mat <- cov(dim_rel_scaled)
m1_cov_mat[upper.tri(m1_cov_mat)] <- NA #Means to asign NA to the elements above the diagonal
m1_cov_mat
```


## Define a model with one latent factor and estimate it
```{r}
hw10_m1 <- 'single_factor =~ Communal.Sharing + Endurance + Activity.Intensity + Activeness + Uniqueness + Attachment + Love.Expression + Mating + Expected.Reciprocity + Intimacy + Importance.for.individuals.involved +  Occupational + Strategic + Formality.and.Regulation + Spatial.Distance + Synchronicity + Conflict + Coercion + Negotiation + Goods.Exchange + Concreteness'

hw10_m1_fit <- cfa(hw10_m1, dim_rel_scaled)
```

## Summary of fit information
```{r}
fitMeasures(hw10_m1_fit, c("npar", "chisq", "df", "pvalue", "cfi", "rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "rmsea.pvalue", "BIC", "AIC"))
```

## Return Coefficients for unstandardized and standardized models
```{r}
hw10_m1_fit_c <- parameterEstimates(hw10_m1_fit)
                   
hw10_m1_fit_sc <-parameterEstimates(hw10_m1_fit, standardized = T)
hw10_m1_fit_sc
```


## Write model for the 3 factor model
```{r}
socio_indicators <- 'Socioemotional =~ Communal.Sharing + Endurance + Activity.Intensity + Activeness + Uniqueness + Attachment + Love.Expression + Mating + Expected.Reciprocity + Intimacy'

exc_indicators <- 'Exchange =~ Importance.for.individuals.involved +  Occupational + Strategic + Formality.and.Regulation + Negotiation + Goods.Exchange + Concreteness'

valence_indicators <- 'Valence.Evaluation =~ Spatial.Distance + Synchronicity + Conflict + Coercion'

hw10_m2 <- paste(socio_indicators, exc_indicators, valence_indicators, sep = "\n")
cat(hw10_m2)
```


## write model for the 4 factor model
### Divide up exchange into information exchange and money exchange
```{r}
exc_opp_indicators <- 'Information.Exchange =~ Importance.for.individuals.involved +  Occupational + Strategic + Formality.and.Regulation'

mon_ex_indicators <- 'Money.Exchange =~ Negotiation + Goods.Exchange + Concreteness'

hw10_m3 <- paste(socio_indicators, exc_opp_indicators, valence_indicators, mon_ex_indicators, sep = "\n")
cat(hw10_m3)
```


## Estimate 3 factor model
```{r}
hw10_m2_fit <- cfa(hw10_m2, dim_rel_scaled, mimic = "Mplus")
```

## Summary of fit information for 3 factor model
```{r}
hw10_m2_fit_sum <- fitMeasures(hw10_m2_fit, c("npar", "chisq", "df", "pvalue", "cfi", "rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "rmsea.pvalue", "BIC", "AIC"))
```

## Estimate 4 factor model
```{r}
hw10_m3_fit <- cfa(hw10_m3, dim_rel_scaled, mimic = "Mplus")
```

## Summary of fit information for 4 factor model
```{r}
hw10_m3_fit_sum <- fitMeasures(hw10_m3_fit, c("npar", "chisq", "df", "pvalue", "cfi", "rmsea", "rmsea.ci.lower", "rmsea.ci.upper", "rmsea.pvalue", "BIC", "AIC"))
```


## Put fit information together in a more helpful format
```{r}
round(cbind(hw10_m2_fit_sum, hw10_m3_fit_sum), 3)
```


## Return Coefficients for each model
```{r}
hw10_m2_fit_sc <-parameterEstimates(hw10_m2_fit, standardized = T)
hw10_m3_fit_sc <-parameterEstimates(hw10_m3_fit, standardized = T)

hw10_m2_m3_all_param <- cbind.fill(subset(hw10_m2_fit_sc, !(op == "~1"))[c(1:4,11)], subset(hw10_m3_fit_sc, !(op == "~1"))[c(1:4,11)], fill = "")
write.csv(hw10_m2_m3_all_param, 'Model 2 & Model 3 Parameters.csv')
hw10_m2_m3_all_param
```



# Results
A confirmatory factory analysis (CFA) was done to see if multiple dimensions that can describe social relationships, can be summarized as a single latent factor. 



