---
title: "Assignment 2 - Path Analysis"
output: html_notebook
---

```{r}
#library(MVN)
library(lavaan)
library(haven)
library(semPlot)
```

```{r}
setwd('~/Google_Drive/courses/Structural_Equation_Modeling/assignments/assignment_02/')
getwd()
```

# Replicating Path Analysis from Sexton et al., (2003)
[link to article](https://www.tandfonline.com/doi/abs/10.1080/16506070302321)

## Create a correlation matrix and a covariance matrix
```{r}
# Input information to create the matrix of information; notice that it is a square matrix with 1s on the diagonal
hw2_cor_mat <- matrix(c(1, .566, .534, .580, .414, .374, .665,
                        .566, 1, .487, .549, .583, .478, .535, 
                        .534, .487, 1, .429, .273, .284, .569, 
                        .580, .549, .429, 1, .492, .443, .554, 
                        .414, .583, .273, .492, 1, .411, .546,
                        .374, .478, .284, .443, .411, 1, .332,
                        .665, .535, .569, .554, .546, .332, 1), 
                    nrow = 7, ncol = 7)

# A numeric vector of the SDs of the variables
var_sds <- c(9.40, 23.34, 18.05, 9.11, 13.77, 15.20, 14.89)

# A list of the variable names for the matrix
var_names = c("NEO_N", "ASI_R", "IUS", "BAI", "IAS", "PI_WSUR", "PSWQ")

# Putting all of the information together to make a covariance matrix
hw2_cov <- cor2cov(hw2_cor_mat, var_sds, names = var_names)
hw2_cov
```


## Replicate regression analysis
I was having trouble replicating the entire hierarchical analysis, so I decided to just do one arm of the analysis.
```{r}
# Define the model of interest
hw2_m1 <- 'BAI ~ ASI_R + NEO_N + IUS
           ASI_R ~ NEO_N
           IUS ~ NEO_N'

# Estimate the model
hw2_m1_fit <- sem(hw2_m1, sample.cov = hw2_cov, sample.nobs = 91)

# Examine initial model fit information
hw2_m1_fit

# Examine more information
summary(hw2_m1_fit)
```
There was a failure to completely replicated the findings in Sexton et al. (2003). The betas of the replication analysis are very different from the betas in the hierarchical path analysis from the original paper. What is consistent, however, is the significant paths. Panic symptoms (BAI) is predicted by anxiety sensitivity (ASI_R) and neuroticism (NEO_N), but not by intolerance of uncertainty (IUS). This could be due to missing variance from the item level data for participants. The individual differences are not fully captured by the univariate standard deviation, nor the correlation matrix or resulting covariance matrix. 



# Path Analysis on Observed Data

```{r}
## Load dataframe of raw data
hw2_df <- read.csv('../unrestricted_hspopal_9_9_2019_22_40_57.csv')

```

## Information about the dataset
This dataset is the Human Connectome Project data for healthy young adults. I will be looking at neuroticism from the NEO-PI, a measure of theory of mind, a measure of friendship, and a measure of angry affect.

# Estimate using FIML
```{r}
hw2_m2 <- 'NEOFAC_O ~ Social_Task_Perc_TOM + Friendship_Unadj + AngAffect_Unadj
           NEOFAC_N ~ Social_Task_Perc_TOM + Friendship_Unadj + AngAffect_Unadj
           Friendship_Unadj ~~ AngAffect_Unadj
           Friendship_Unadj ~~ Social_Task_Perc_TOM
           Social_Task_Perc_TOM ~~ AngAffect_Unadj'

# Estimation of FIML model
hw2_m2_fit <- sem(hw2_m2, hw2_df, missing = "fiml")

## Model Information FIML model
summary(hw2_m2_fit)
```

## Figure of model
```{r}
semPaths(hw2_m2_fit, "path", whatLabels = "est", 
         edge.label.cex = 1, rotation = 2, intercepts = FALSE, residuals = TRUE, curve = 1, 
         curvature = 3, nCharNodes = 8, sizeMan = 6, sizeMan2 = 3, optimizeLatRes = TRUE, edge.color = "#000000")
```

# Estimate without using FIML
```{r}
hw2_m3 <- 'NEOFAC_O ~ Social_Task_Perc_TOM + Friendship_Unadj + AngAffect_Unadj
           NEOFAC_N ~ Social_Task_Perc_TOM + Friendship_Unadj + AngAffect_Unadj'

# Estimation of FIML model
hw2_m3_fit <- sem(hw2_m3, hw2_df)

## Model Information FIML model
summary(hw2_m3_fit)
```

Running this analysis with FIML and without resulted in slightly different betas. Some predictors were also not significant when FIML was not done.

## Figure of model
```{r}
semPaths(hw2_m3_fit, "path", whatLabels = "est", 
         edge.label.cex = 1, rotation = 2, intercepts = FALSE, residuals = TRUE, curve = 1, 
         curvature = 3, nCharNodes = 8, sizeMan = 6, sizeMan2 = 3, optimizeLatRes = TRUE, edge.color = "#000000")
```

