---
title: "Assignment 4"
output: html_notebook
---

```{r}
library(tidyverse)
library(haven)
library(psych)
library(reghelper)
library(sjPlot)
```

Set working directory and import data
```{r}

setwd('~/Google_Drive/courses/Multivariate/homework/assignment_04/')
getwd()

hcp_behav <- read.csv("~/Google_Drive/courses/Multivariate/homework/assignment_01/unrestricted_hspopal_9_9_2019_22_40_57.csv", row.names = 1, header = TRUE)
```
```{r}
hcp_behav$Gender_group[hcp_behav$Gender == 'F'] <- 1
hcp_behav$Gender_group[hcp_behav$Gender == 'M'] <- 2
#hcp_behav$Age_group[hcp_behav$Age == '31-35'] <- 3
#hcp_behav$Age_group[hcp_behav$Age == '36+'] <- 4
```

### Compute centered variable values
```{r}
hcp_behav <- mutate(hcp_behav, Gender_group_c = Gender_group - mean(Gender_group, na.rm = T),
                              Friendship_Unadj_c = Friendship_Unadj - mean(Friendship_Unadj, na.rm = T),
                             genderXfriendship = Gender_group_c*Friendship_Unadj_c)
describe(hcp_behav[c('Gender_group','Friendship_Unadj',
                     'Gender_group_c','Friendship_Unadj_c','genderXfriendship')])

```

### Run regression model for centered variables
```{r}
h4_m1 <- lm(LifeSatisf_Unadj ~ Gender_group_c + Friendship_Unadj_c, hcp_behav)
summary(h4_m1)
```
### Run regression model for centered variables and interaction
```{r}
h4_m2 <- lm(LifeSatisf_Unadj ~ Gender_group_c + Friendship_Unadj_c + genderXfriendship, hcp_behav)
summary(h4_m2)
```

# Compute possibc variables for +/- SD around the mean
Center the data so that the person who was at 0.97, is now at zero.
```{r}
hcp_behav <- mutate(hcp_behav, AngAffect_m1 = AngAffect_Unadj_c + sd(AngAffect_Unadj_c, 
                                                                    na.rm = T),
                            AngAffect_p1 = AngAffect_Unadj_c - sd(AngAffect_Unadj_c, 
                                                                 na.rm = T),
                             angaffect_m1Xfrnd = AngAffect_m1*Friendship_Unadj_c,
                             angaffect_p1Xfrnd = AngAffect_p1*Friendship_Unadj_c)

describe(hcp_behav[c("AngAffect_Unadj", "AngAffect_Unadj_c", "AngAffect_m1", 
                     "AngAffect_p1")])
```

# Model for simple slopes at 1 SD below mean of positive sibling relationships
```{r}
h4_m3 <- lm(LifeSatisf_Unadj ~ AngAffect_m1 + Friendship_Unadj_c + angaffect_m1Xfrnd,
            hcp_behav)
summary(h4_m3)
```

# Model for simple slopes at 1 SD above mean of positive sibling relationships

```{r}
h4_m4 <- lm(LifeSatisf_Unadj ~ AngAffect_p1 + Friendship_Unadj_c + angaffect_p1Xfrnd, 
            hcp_behav)
summary(h4_m4)
```
```{r}
anova(h4_m3, h4_m4)
```


# Using some shortcuts
## For continuous X continuous interaction
```{r}
h4_m5 <- lm(LifeSatisf_Unadj ~ AngAffect_Unadj_c*Friendship_Unadj_c, hcp_behav)
summary(h4_m5)

h4_m5_ss <- simple_slopes(h4_m5)
h4_m5_ss
```

# Continuous X Continuous Plot
```{r}
h4_m5_p <- plot_model(h4_m5, type = "int")
h4_m5_p

# Also plot_model(cl5_m11, type = "diag") for regression diagnostics that are clear!
plot_model(h4_m5, type = "diag")
```












