---
title: "Assignment 6 - Measurement invariance for well-being across sex"
output: html_notebook
---

```{r}
library(psych)
library(MVN)
library(lavaan)
library(lavaanPlot)
library(tidyverse)
library(apaTables)
library(patchwork)
library(semTools)
library(corrplot)
```

```{r}
setwd('~/Google_Drive/courses/Structural_Equation_Modeling/assignments/assignment_06/')
getwd()
```


Import data
```{r}
cl7_df1 <- read.delim('s6.9 Class 6. Mult Grp - Simulation.dat',
                      header = F, sep = "")

names(cl7_df1) <- c("SP_COMP", "SP_ABL1", "SP_ABL2", "SP_INTER", "WB1", "WB2", "WB3", "WB4", "Female")
```

```{r}
corr_mtx <- apa.cor.table(cl7_df1)
corr_mtx
```


### Rather than creating two separate dataframes, we'll work with subsetting the dataframe and keep it in tact
```{r}
trim_fit <- function(inp, outname) {
  #Multiple steps to get smaller amount of information
  fit_res <- data.frame(fitMeasures(inp))
  fit_res <- round(fit_res, 3)
  fit_res$index <- row.names(fit_res)
  #rownames(fit_res) <- c()
  
#  Retaining only the commonly used indices; the last '[c(2,1)]' just changes the 
#  order of the columns
  out <- fit_res[fit_res$index %in% c("npar", "chisq", "df", "pvalue",
                                          "rmsea", "rmsea.ci.lower", "rmsea.ci.upper",
                                          "rmsea.pvalue", "cfi", "tli", "aic", "bic", "bic2"),][c(2,1)]
  names(out) <- c("index", "value")
  assign(deparse(substitute(outname)), out, envir=.GlobalEnv)
}
```

# Well-Being
## Configural invariance model
```{r}
Wellb_mod_mg <- 'Wellb =~ NA*WB1 + WB2 + WB3 + WB4
                   Wellb ~~ 1*Wellb'

Wellb_Configural_res <- sem(Wellb_mod_mg, cl7_df1, group = "Female")
Wellb_Configural_res
```
### Model fit information
```{r}
trim_fit(Wellb_Configural_res, Wellb_Configural_res_fit)

#View model fit information
Wellb_Configural_res_fit
```

### Parameter Estimates
```{r}
parameterestimates(Wellb_Configural_res, ci = F, standardized = T)
```
Well being latent factor is fixed at both places. By default, latent variable means are fixed at 0.

### lavaan plot
```{r}
wellb_conf_fig <- lavaanPlot(model = Wellb_Configural_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
wellb_conf_fig
```

## Metric invariance model
```{r}
Wellb_mod_mg_met <- 'Wellb =~ c(NA, NA)*WB1 + c(l1, l1)*WB1 + c(l2, l2)*WB2 + c(l3, l3)*WB3 + c(l4, l4)*WB4
                   Wellb ~~ c(1, NA)*Wellb'

Wellb_Metric_res <- sem(Wellb_mod_mg_met, cl7_df1, group = "Female")
```
### Model fit information
```{r}
trim_fit(Wellb_Metric_res, Wellb_Metric_res_fit)

#View model fit information
Wellb_Metric_res_fit
```
### Parameter estimates
```{r}
parameterestimates(Wellb_Metric_res, ci = F, standardized = T)
```

### lavaan plot
```{r}
wellb_metr_fig <- lavaanPlot(model = Wellb_Metric_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
wellb_metr_fig
```


## Scalar invariance model
```{r}
Wellb_mod_mg_scalar <- 'Wellb =~ c(NA, NA)*WB1 + c(l1, l1)*WB1 + c(l2, l2)*WB2 + c(l3, l3)*WB3 + c(l4, l4)*WB4
                   Wellb ~~ c(1, NA)*Wellb
                   WB1 ~ c(i1, i1)*1
                   WB2 ~ c(i2, i2)*1
                   WB3 ~ c(i3, i3)*1
                   WB4 ~ c(i4, i4)*1
                   Wellb ~ c(0, NA)*1'

Wellb_Scalar_res <- sem(Wellb_mod_mg_scalar, cl7_df1, group = "Female")
```
### Model fit information
```{r}
trim_fit(Wellb_Scalar_res, Wellb_Scalar_res_fit)

#View model fit information
Wellb_Scalar_res_fit
```
### Parameter Estimates
```{r}
parameterestimates(Wellb_Scalar_res, ci = F, standardized = T)
```

### lavaan plot
```{r}
wellb_scal_fig <- lavaanPlot(model = Wellb_Scalar_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
wellb_scal_fig
```

## Pulling all model fit together
```{r}
all_Wellb_fit <- cbind(Wellb_Configural_res_fit, Wellb_Metric_res_fit[2], Wellb_Scalar_res_fit[2])
names(all_Wellb_fit) <- c("Index", "Configural Model", "Metric Model", "Scalar Model")
write.table(all_Wellb_fit, "models_fit.csv", sep=",")
all_Wellb_fit
```



# Sports
## Configural invariance model
```{r}
cl7_Sports_mod_mg <- 'Sports =~ NA*SP_COMP + SP_ABL1 + SP_ABL2 + SP_INTER
                   Sports ~~ 1*Sports'

cl7_Sports_Configural_res <- sem(cl7_Sports_mod_mg, cl7_df1, group = "Female")
```

### Model fit information
```{r}
trim_fit(cl7_Sports_Configural_res, cl7_Sports_Configural_res_fit)

#View model fit information
cl7_Sports_Configural_res_fit
```

### Parameter Estimates
```{r}
parameterestimates(cl7_Sports_Configural_res, ci = F, standardized = T)
```

### lavaan plot
```{r}
sports_conf_fig <- lavaanPlot(model = cl7_Sports_Configural_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
sports_conf_fig
```

## Metric invariance model
```{r}
cl7_Sports_mod_mg_met <- 'Sports =~ c(NA, NA)*SP_COMP + c(l1, l1)*SP_COMP + c(l2, l2)*SP_ABL1 + c(l3, l3)*SP_ABL2 + c(l4, l4)*SP_INTER
                   Sports ~~ c(1, NA)*Sports'

cl7_Sports_Metric_res <- sem(cl7_Sports_mod_mg_met, cl7_df1, group = "Female")
```

### Model fit information
```{r}
trim_fit(cl7_Sports_Metric_res, cl7_Sports_Metric_res_fit)

#View model fit information
cl7_Sports_Metric_res_fit
```

### Parameter estimates
```{r}
parameterestimates(cl7_Sports_Metric_res, ci = F, standardized = T)
```

### lavaan plot
```{r}
sports_metr_fig <- lavaanPlot(model = cl7_Sports_Metric_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
sports_metr_fig
```

## Scalar invariance model
```{r}
cl7_Sports_mod_mg_scalar <- 'Sports =~ c(NA, NA)*SP_COMP + c(l1, l1)*SP_COMP + c(l2, l2)*SP_ABL1 + c(l3, l3)*SP_ABL2 + c(l4, l4)*SP_INTER
                   Sports ~~ c(1, NA)*Sports
                   SP_COMP ~ c(i1, i1)*1
                   SP_ABL1 ~ c(i2, i2)*1
                   SP_ABL2 ~ c(i3, i3)*1
                   SP_INTER ~ c(i4, i4)*1
                   Sports ~ c(0, NA)*1'

cl7_Sports_Scalar_res <- sem(cl7_Sports_mod_mg_scalar, cl7_df1, group = "Female")
```
### Model fit information
```{r}
trim_fit(cl7_Sports_Scalar_res, cl7_Sports_Scalar_res_fit)

#View model fit information
cl7_Sports_Scalar_res_fit
```
### Parameter Estimates
```{r}
parameterestimates(cl7_Sports_Scalar_res, ci = F, standardized = T)
```

### lavaan plot
```{r}
sports_scal_fig <- lavaanPlot(model = cl7_Sports_Scalar_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
sports_scal_fig
```

## Pulling all model fit together
```{r}
all_Sports_fit <- cbind(cl7_Sports_Configural_res_fit, cl7_Sports_Metric_res_fit[2], cl7_Sports_Scalar_res_fit[2])
names(all_Sports_fit) <- c("Index", "Configural Model", "Metric Model", "Scalar Model")
write.table(all_Sports_fit, "models_fit_sports.csv", sep=",")
all_Sports_fit
```

## Partial Metric Invariance
```{r}
cl7_Sports_mod_mg_p_met <- 'Sports =~ c(NA, NA)*SP_COMP + c(l1, l1)*SP_COMP + c(l2, l2)*SP_ABL1 + c(l3, l3)*SP_ABL2 + SP_INTER
                   Sports ~~ c(1, NA)*Sports'

cl7_Sports_Part_Metric_res <- sem(cl7_Sports_mod_mg_p_met, cl7_df1, group = "Female")
```

### Model fit information
```{r}
trim_fit(cl7_Sports_Part_Metric_res, cl7_Sports_Part_Metric_res_fit)

#View model fit information
cl7_Sports_Part_Metric_res_fit
```

### Parameter estimates
```{r}
parameterestimates(cl7_Sports_Part_Metric_res, ci = F, standardized = T)
```

### lavaan plot
```{r}
sports_pmetr_fig <- lavaanPlot(model = cl7_Sports_Part_Metric_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
sports_pmetr_fig
```

## Partial Scalar Invariance
```{r}
cl7_Sports_mod_mg_p_scalar <- 'Sports =~ c(NA, NA)*SP_COMP + c(l1, l1)*SP_COMP + c(l2, l2)*SP_ABL1 + c(l3, l3)*SP_ABL2 + SP_INTER
                   Sports ~~ c(1, NA)*Sports
                   SP_COMP ~ c(i1, i1)*1
                   SP_ABL1 ~ c(i2, i2)*1
                   SP_ABL2 ~ c(i3, i3)*1
                   SP_INTER ~ 1
                   Sports ~ c(0, NA)*1'

cl7_Sports_Part_scalar_res <- sem(cl7_Sports_mod_mg_p_scalar, cl7_df1, group = "Female")
```

### Model fit information
```{r}
trim_fit(cl7_Sports_Part_scalar_res, cl7_Sports_Part_scalar_res_fit)

#View model fit information
cl7_Sports_Part_scalar_res_fit
```

### Parameter Estimates
```{r}
parameterestimates(cl7_Sports_Part_scalar_res, ci = F, standardized = T)
```

## Pulling all model fit together; With Partial invariance
```{r}
all_Sports_fit_with_partial <- cbind(cl7_Sports_Configural_res_fit, cl7_Sports_Metric_res_fit[2], cl7_Sports_Part_Metric_res_fit[2], 
                        cl7_Sports_Scalar_res_fit[2], cl7_Sports_Part_scalar_res_fit[2])
names(all_Sports_fit_with_partial) <- c("Index", "Configural Model", "Metric Model", "Partial Metric Model", "Scalar Model", "Partial Scalar Model")
write.table(all_Sports_fit_with_partial, "models_fit_sports_partial.csv", sep=",")

all_Sports_fit_with_partial
```


# Moderation
```{r}
mod_sports_wellb <- 'Sports =~ c(NA, NA)*SP_COMP + c(l1, l1)*SP_COMP + c(l2, l2)*SP_ABL1 + c(l3, l3)*SP_ABL2 + SP_INTER
                   Wellb =~ c(NA, NA)*WB1 + c(l1, l1)*WB1 + c(l2, l2)*WB2 + c(l3, l3)*WB3 + c(l4, l4)*WB4
                   
                   Sports ~~ c(1, NA)*Sports
                   SP_COMP ~ c(i1, i1)*1
                   SP_ABL1 ~ c(i2, i2)*1
                   SP_ABL2 ~ c(i3, i3)*1
                   SP_INTER ~ 1
                   Sports ~ c(0, NA)*1

                   Wellb ~~ c(1, NA)*Wellb
                   WB1 ~ c(i1, i1)*1
                   WB2 ~ c(i2, i2)*1
                   WB3 ~ c(i3, i3)*1
                   WB4 ~ c(i4, i4)*1
                   Wellb ~ c(0, NA)*1
                   
                   Wellb ~ Sports'

mod_sports_wellb_res <- sem(mod_sports_wellb, cl7_df1, group="Female")
```

### Model fit information
```{r}
trim_fit(mod_sports_wellb_res, mod_sports_wellb_res_fit)

#View model fit information
mod_sports_wellb_res_fit
```


### Parameter Estimates
```{r}
parameterestimates(mod_sports_wellb_res, ci = F, standardized = T)
```

### lavaan plot
```{r}
sports_wellb_mod_fig <- lavaanPlot(model = mod_sports_wellb_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
sports_wellb_mod_fig
```


## Moderation with contrained factors

```{r}
mod_sports_wellb_con <- "#partial scalar invariance model for Sports factor
                  Sports =~ c(NA, NA)*SP_COMP + c(l1, l1)*SP_COMP + c(l2, l2)*SP_ABL1 + c(l3, l3)*SP_ABL2 + SP_INTER
                  Sports ~~ c(1, NA)*Sports
                  SP_COMP ~ c(i1, i1)*1
                  SP_ABL1 ~ c(i2, i2)*1
                  SP_ABL2 ~ c(i3, i3)*1
                  SP_INTER ~ 1
                  Sports ~ c(0, NA)*1

                  #full scalar invariance model for Well-Being factor
                  WellBeing =~ c(NA, NA)*WB1 + c(l4, l4)*WB1 + c(l5, l5)*WB2 + c(l6, l6)*WB3 + c(l7, l7)*WB4
                  WellBeing ~~ c(1, NA)*WellBeing
                  WB1 ~ c(i4, i4)*1
                  WB2 ~ c(i5, i5)*1
                  WB3 ~ c(i6, i6)*1
                  WB4 ~ c(i7, i7)*1
                  WellBeing ~ c(0, NA)*1

                  #latent variable regression
                  WellBeing ~ c(l8, l8)*Sports" #constrain male and females loadings to be equal
mod_sports_wellb_con_res <- sem(mod_sports_wellb_con, cl7_df1, group="Female")
```

### Model fit information
```{r}
trim_fit(mod_sports_wellb_con_res, mod_sports_wellb_con_res_fit)

#View model fit information
mod_sports_wellb_con_res_fit
```

### Pulling all model fit together
```{r}
all_mod_fit <- cbind(mod_sports_wellb_res_fit, mod_sports_wellb_con_res_fit[2])
names(all_mod_fit) <- c("Index", "Unconstrained", "Constrained")
write.table(all_mod_fit, "models_fit_moderation_con.csv", sep=",")

all_mod_fit
```


### Parameter Estimates
```{r}
parameterestimates(mod_sports_wellb_con_res, ci = F, standardized = T)
```


### lavaan plot
```{r}
sports_wellb_mod_con_fig <- lavaanPlot(model = mod_sports_wellb_con_res, 
           node_options = list(shape = "box", fontname = "Helvetica"), 
           edge_options = list(color = "grey"), 
           graph_options = list(layout = "dot", rankdir = "TB"),
           coefs = T, stand = F, sig = 1.00, covs = T, stars = c("regress", "latent"))
sports_wellb_mod_con_fig
```


## Chi-Square Difference Test
```{r}
anova(mod_sports_wellb_res, mod_sports_wellb_con_res)
```


