---
title: "Assignment 6"
output: html_notebook
---

```{r}
library(tidyverse)
library(haven)
library(psych)
library(lavaan)
#library(semPlot)
```

# Note: I am having trouble installing the semPlot package, and so did not include any of that code in this script

```{r}
setwd('~/Google_Drive/courses/Multivariate/homework/assignment_06')
getwd()

h6_df <- read_sav('Assignment 6 Conditional Process Data.sav')
names(h6_df)
```

## Estimate the first model with just the X, Y, and mediator
```{r}
h6_mod1 <-   'aggress ~ b*cogcon + c*sleep + provoc
              cogcon ~ a*sleep + provoc
              #Indirect effect
                ab := a*b
                #Total effect
                tot_eff := c + (a*b)'


h6_m1 <- sem(h6_mod1, h6_df)
summary(h6_m1)
```


### Bootstrap the CIs
```{r}
parameterEstimates(h6_m1, standardized = T, boot.ci.type = "bca.simple")
```
The condition indirect effect of provocation is significant as the p < .05 and the CI does not include 0

## Modeling the interaction of sleep and provocation and the indirect effects

### create centered variables and interaction terms
```{r}
h6_df$sleepc = h6_df$sleep - mean(h6_df$sleep, na.rm = T)
h6_df$cogconc = h6_df$cogcon - mean(h6_df$cogcon, na.rm = T)
h6_df$provocc = h6_df$provoc - mean(h6_df$provoc, na.rm = T)
h6_df$sleep_provoc = h6_df$provocc*h6_df$sleepc

describe(h6_df)
```

### Adding the interaction only on the a path
```{r}
h6_mod2 <-   'cogconc ~ a1*sleepc + a2* provocc + a3*sleep_provoc
              aggress ~ b1*cogconc + c*sleepc
              #Indirect effect
                INDHighProv := b1*(a1 + a3*-.528)
                INDLowProv := b1*(a1 + a3*0.472)
              #Total effect
                TOTHighProv := INDHighProv + c
                TOTLowProv := INDLowProv + c
                IndexMM := INDHighProv - INDLowProv'

h6_m2 <- sem(h6_mod2, h6_df)
summary(h6_m2)
```
### Bootstrap the CIs
```{r}
parameterEstimates(h6_m2, standardized = T, boot.ci.type = "bca.simple")
```


## Estimating the conditional indirect effect for with high provocation
```{r}
h6_df$provoc_high = h6_df$provocc + .528
h6_df$sleep_provoc_high = h6_df$provoc_high*h6_df$sleepc

h6_mod3 <-   'cogconc ~ a1*sleepc + a2* provoc_high + a3*sleep_provoc_high
              aggress ~ b1*cogconc + c*sleepc
              #Indirect effect
                INDHighProv := a1*b1'

h6_m3 <- sem(h6_mod3, h6_df)
summary(h6_m3)

parameterEstimates(h6_m3, standardized = T, boot.ci.type = "bca.simple")
```

## Estimating the conditional indirect effect for with low provocation
```{r}
h6_df$provoc_low = h6_df$provocc - .472
h6_df$sleep_provoc_low = h6_df$provoc_low*h6_df$sleepc

h6_mod4 <-   'cogconc ~ a1*sleepc + a2* provoc_low + a3*sleep_provoc_low
              aggress ~ b1*cogconc + c*sleepc
              #Indirect effect
                INDlowProv := a1*b1'

h6_m4 <- sem(h6_mod4, h6_df)
summary(h6_m4)

parameterEstimates(h6_m4, standardized = T, boot.ci.type = "bca.simple")
```







