---
title: "Assignment 7 - Logistic Regression"
output: html_notebook
---


```{r}
library(haven)
library(psych)
library(tidyverse)
library(effects)
library(reghelper)
```

```{r}
setwd('~/Google_Drive/courses/Multivariate/homework/assignment_07/')
getwd()

hcp_behav <- read.csv("../unrestricted_hspopal_9_9_2019_22_40_57.csv", row.names = 1, header = TRUE)

#Simple Descriptive Statistics
describe(hcp_behav)
str(hcp_behav)
```

# Logistic Regression
### Estimate model & View results
```{r}
h7m1 <- glm(Gender ~ LifeSatisf_Unadj + Friendship_Unadj, hcp_behav, family = "binomial")
summary(h7m1)
```

### Obtain predicted probability values then plot them; defaults effects plot in base R scales information oddly
```{r}
h7m1_pred_x1 <- data.frame(effect("LifeSatisf_Unadj", h7m1, xlevels=list(LifeSatisf_Unadj=0:75)))

h7m1_pred_x1_pl <- ggplot(h7m1_pred_x1, aes(x=LifeSatisf_Unadj, y=fit)) + geom_point() + geom_smooth() + geom_ribbon(aes(ymin=lower, ymax=upper), alpha = .1)
h7m1_pred_x1_pl

h7m1_pred_x2 <- data.frame(effect("Friendship_Unadj", h7m1, xlevels=list(Friendship_Unadj=0:70)))

h7m1_pred_x2_pl <- ggplot(h7m1_pred_x2, aes(x=Friendship_Unadj, y=fit)) + geom_point() + geom_smooth() + geom_ribbon(aes(ymin=lower, ymax=upper), alpha = .1)
h7m1_pred_x2_pl
```


# Moderation Analyses
### Centering and defining a factor variable
```{r}
hcp_behav$LifeSatisf_Unadjc <- hcp_behav$LifeSatisf_Unadj - mean(hcp_behav$LifeSatisf_Unadj, na.rm = TRUE)
hcp_behav$Friendship_Unadjc <- hcp_behav$Friendship_Unadj - mean(hcp_behav$Friendship_Unadj, na.rm = TRUE)
hcp_behav$AngAffect_Unadjc <- hcp_behav$AngAffect_Unadj - mean(hcp_behav$AngAffect_Unadj, na.rm = TRUE)
```

### Specifying and estimating the model
This model has our three main effects, and we are including all three of the interaciton effects in the same model. We have also changed it from a continuous to a binary model.
```{r}
h7m2 <- glm(Gender ~ LifeSatisf_Unadjc + Friendship_Unadjc + AngAffect_Unadjc + LifeSatisf_Unadjc*AngAffect_Unadjc + Friendship_Unadjc*AngAffect_Unadjc + LifeSatisf_Unadjc*Friendship_Unadjc, hcp_behav, family = "binomial")
summary(h7m2)
```

```{r}
# Create dataframe object for creating the ORs
h7m2c <- data.frame(coefficients(summary(h7m2)))

#Compute ORs and CIs around the point estimates
h7m2c$OR <- round(exp(h7m2c$Estimate), 2)
h7m2c$LL <- round(exp(h7m2c$Estimate -1.96*h7m2c$Std..Error), 2)
h7m2c$UL <- round(exp(h7m2c$Estimate +1.96*h7m2c$Std..Error), 2)
#Concatenate OR & CIs into a single columnM
h7m2c$OR_CI <- paste0(h7m2c$OR," (", h7m2c$LL,"-",h7m2c$UL,")")
#cl8m2c$var <- row.names(cl8m2c)
h7m2c
#Write results into a csv file
write.csv(h7m2c, 'assignment_07_m2_results.csv')
```

## Create effects object for the interaction term
```{r}
h7m2_pred <- effect('LifeSatisf_Unadjc*AngAffect_Unadjc', h7m2)

#Using base R to make a very simple plot
h7m2_pred_pl <- plot(h7m2_pred, ask=FALSE, rescale.axis = F)
h7m2_pred_pl
```

## Obtain Simple slopes for significant interaction
```{r}
h7m2_ss <- simple_slopes(h7m2)
h7m2_ss
```





