---
title: "Assignment 3"
output: html_notebook
---
### Load necessary packages
```{r}

library(haven)
library(tidyverse)
library(psych)
library(stargazer)

setwd("~/Google_Drive/courses/Multivariate/homework/assignment_03//")
getwd()
```


# Read in data
```{r}
h3_data <- read_sav('Assignment 3 Data.sav')
names(h3_data)
str(h3_data)
```

# Recode intervention group variable into a four-level categorical variable
```{r}
h3_data$group_cat[h3_data$GROUP == 1]  <- "both"
h3_data$group_cat[h3_data$GROUP == 2]  <- "individual"
h3_data$group_cat[h3_data$GROUP == 3] <- "group"
h3_data$group_cat[h3_data$GROUP == 4] <- "no intervention"

# Initial comparison using ANOVA
h3_mod1 <- aov(POSTREAD ~ group_cat, h3_data)
summary(h3_mod1)
```


# Pairwise comparisone using Tukey's HSD Test
```{r}
TukeyHSD(h3_mod1)
```

# Comparison using LM function
```{r}
h3_mod2 <- lm(POSTREAD ~ group_cat, h3_data)
summary(h3_mod2)
```

# Creating dummy codes for "under the hood" regression models; reference category is no intervention
```{r}
h3_data$ps_gvb[h3_data$group_cat == "individual" | h3_data$group_cat == "no intervention" | h3_data$group_cat == "both"] <- 0
h3_data$ps_gvb[h3_data$group_cat == "group"] <- 1
h3_data$ps_ivb[h3_data$group_cat == "group" | h3_data$group_cat == "no intervention" | h3_data$group_cat == "both"]<- 0
h3_data$ps_ivb[h3_data$group_cat == "individual"] <- 1
h3_data$ps_nvb[h3_data$group_cat == "group" | h3_data$group_cat == "individual" | h3_data$group_cat == "both"] <- 0
h3_data$ps_nvb[h3_data$group_cat == "no intervention"] <- 1

#LM With Dummy Codes
h3_mod3 <- lm(POSTREAD ~ ps_gvb + ps_ivb + ps_nvb, h3_data)
summary(h3_mod3)

h3_mod4 <- lm(POSTREAD ~ ps_ivb + ps_gvb + ps_nvb, h3_data)
summary(h3_mod4)

h3_mod5 <- lm(POSTREAD ~ ps_nvb + ps_gvb + ps_ivb, h3_data)
summary(h3_mod5)
```

# Examine the influence of treatment condition accounting for pre-treatment reading, participant age, and intellectual functioning scores
```{r}
mr_h3.m1 <- lm(POSTREAD ~ group_cat + PREREAD + AGE + IQ, h3_data)
summary(mr_h3.m1)

stargazer(mr_h3.m1, type = "text", title = "Intervention Group and Reading Test Scores", 
          single.row = T,  out="question2_output.htm")
```


